Oregon 2013 Medication for Asthma Inhalers

### 2017-07-12 r_scripts/analysis/oregon_medication.Rmd
Update oregon_medication.Rmd

```{r regression model q, message=FALSE, echo=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")

@@

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

```{r, message=FALSE, echo=FALSE}
# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=6)
colnames(fit_mat) <- c("outcome", "exposure", "fit_mix", "df", "aic", "fit_fix")

@@

```{r compare mixed and fixed, message=FALSE, echo=FALSE}

@@

## ggplot 
  print_plot <- ggplot(wrf_geo_lag_plot,
    aes(x = lag, y = odds_ratio, color = lag), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Geo-smk Lag", 
                       values = c("red", "orange", "yellow", "green", "blue", "#32115C", "purple"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 7, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

@@

@@

@@

>>>>

```{r regression model q, message=FALSE, echo=FALSE, eval=FALSE, warning=FALSE}
# create empty dataframe
rr_df <- data.frame(matrix(nrow = 3, ncol = 5))
colnames(rr_df) <- c("method", "outcome", "rel_risk", "lower_95", "upper_95")

@@

```{r DLM, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

@@

```{r, message=FALSE, echo=FALSE, eval=FALSE}
# create fit matrix
fit_mat <- matrix(NA, nrow=4, ncol=6)
colnames(fit_mat) <- c("outcome", "exposure", "fit_mix", "df", "aic", "fit_fix")

@@

```{r compare mixed and fixed, message=FALSE, echo=FALSE, eval=FALSE}

@@

## ggplot 
  print_plot <- ggplot(wrf_geo_lag_plot,
    aes(x = lag, y = odds_ratio, color = lag), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Geo-smk Lag", 
                       values = rep(c("black"), 7),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 7, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Smoke Estimation Method') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

@@

inhaler_ndc_lag_df <- inhaler_ndc_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)

@@

B <- ns(0:lag_max, df =3, intercept=TRUE)

@@

```{r DLM for casecrossover temp, echo = FALSE, message=FALSE, warning=FALSE}
read_path8 <- paste0("../data_new/county_data/zip_to_county_new.csv")
or_zip_to_county <- read_csv(read_path8)

inhaler_ndc_lag_df <- inhaler_ndc_df %>%
  left_join(or_zip_to_county, by = "ZIPCODE") %>%
  left_join(or_pop, by = "county") %>%
  mutate(day = ifelse(weekdays(date) == "Monday" |
                      weekdays(date) == "Tuesday" |
                      weekdays(date) == "Wednesday"|
                      weekdays(date) == "Thursday"|
                      weekdays(date) == "Friday", "weekday",
               ifelse(weekdays(date) == "Saturday" |
                      weekdays(date) == "Sunday", "weekend", NA))) %>%
  mutate(geo_smk_pm10_zip = geo_smk_pm_zip/10,
         geo_smk_pm10_lag1_zip =geo_smk_pm_lag1_zip/10,
         geo_smk_pm10_lag2_zip =geo_smk_pm_lag2_zip/10,
         geo_smk_pm10_lag3_zip =geo_smk_pm_lag3_zip/10,
         geo_smk_pm10_lag4_zip =geo_smk_pm_lag4_zip/10,
         geo_smk_pm10_lag5_zip =geo_smk_pm_lag5_zip/10,
         geo_smk_pm10_lag6_zip =geo_smk_pm_lag6_zip/10,
         geo_smk_pm10_lag7_zip =geo_smk_pm_lag7_zip/10)

B <- ns(0:lag_max, df =3, intercept=TRUE)

fit <- clogit(outcome ~  inhaler_temp_lag_B + strata(personkey), 
              inhaler_ndc_lag_df)

# dlparms <- grep("inhaler_geo_lag", names(coef(fit)))
dlparms <- c(1,2,3)
DLestimate<- data.frame(estimate=B%*%coef(fit)[dlparms]) 
DLvar <- B%*%vcov(fit)[dlparms,dlparms]%*%t(B)
DLestimate$SE <- sqrt(diag(DLvar))
DLestimate$lower <- DLestimate$estimate - DLestimate$SE * 1.96
DLestimate$upper <- DLestimate$estimate + DLestimate$SE * 1.96

### 3. Unconstrained lag for time series

```{r unconstrained ts lag, message=FALSE, echo=FALSE}
or_ndc_county_ts_df <- or_ndc_county_ts %>%
  select(1:4, pop, rate_per_100k, day, geo_smk_pm_county,
         geo_smk_pm_lag1_county:geo_smk_pm_lag7_county, wrf_temp_county,
         wrf_temp_lag1_county:wrf_temp_lag7_county) %>%
  mutate(geo_smk_pm10_county = geo_smk_pm_county/10,
         geo_smk_pm10_lag1_county =geo_smk_pm_lag1_county/10,
         geo_smk_pm10_lag2_county =geo_smk_pm_lag2_county/10,
         geo_smk_pm10_lag3_county =geo_smk_pm_lag3_county/10,
         geo_smk_pm10_lag4_county =geo_smk_pm_lag4_county/10,
         geo_smk_pm10_lag5_county =geo_smk_pm_lag5_county/10,
         geo_smk_pm10_lag6_county =geo_smk_pm_lag6_county/10,
         geo_smk_pm10_lag7_county =geo_smk_pm_lag7_county/10) 
  # filter(!is.na(geo_smk_pm10_lag7_zip))

# create empty dataframe
rr_df <- data.frame(matrix(nrow = 7, ncol = 5))
colnames(rr_df) <- c("method", "lag", "rel_risk", "lower_95", "upper_95")

# methods
rr_df[,1] <- c("Mixed-Poisson")

# populate outcome column
rr_df[,2] <- c("Lag 1", "Lag 2", "Lag 3", "Lag 4", "Lag 5", "Lag 6", "Lag 7")

# which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") #25
# which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") #17

for (i in 1:7){

  geo <- i + 24
  tem <- i +14

## mixed
smk_est_mp <- tidy(glmer(as.formula(paste0(outcome_list, 
    "~ or_ndc_county_ts_df[[geo]] + or_ndc_county_ts_df[[tem]] + (1|county) + day + offset(log(pop))")), 
    or_ndc_county_ts_df, family="poisson"))[2, ]

rr_df[i,3:5] <- rr_mp
}
rr_df

plot_rr <- ggplot(rr_df, aes(x=lag, y = rel_risk)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower_95, ymax=upper_95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "red") +
  theme(
   panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.text.x = element_text(angle = 90, vjust = .75)) +
  ylab("Relative Risk") +
  xlab("Outcome") +
  ggtitle("Relative Risk for a 10ug/m^3 increase PM2.5 estimated via GWR Smoke")

ggsave("../plot_new/ndc_rr_geo_lag.pdf", plot = plot_rr, 
       width = 12, height = 8, units = "in")

### 4. Distributed lag for time series

```{r DLM ts, echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE}

which(colnames(or_ndc_county_ts_df)=="wrf_temp_lag1_county") # 17
which(colnames(or_ndc_county_ts_df)=="geo_smk_pm10_lag1_county") # 25

geo_ts_lag <- or_ndc_county_ts_df %>%
  select(24:31) # GWR-method
tem_ts_lag <- or_ndc_county_ts_df %>%
  select(16:23) # temp

lag_max <- 7

geo_ts_lag <- as.matrix(geo_ts_lag)
tem_ts_lag <- as.matrix(tem_ts_lag)

# the nrow B should equal ncol(pm)
dim(B)
dim(geo_ts_lag)
dim(tem_ts_lag)

geo_ts_lag_B <- geo_ts_lag%*%B

tem_ts_lag_B <- tem_ts_lag%*%B

## mixed model
mod <-glmer(outcome ~  geo_ts_lag_B + tem_ts_lag_B + day + (1|county) + offset(log(pop)),   
            or_ndc_county_ts_df, family="poisson", control = glmerControl(optimizer = "bobyqa"))
